---
title: "Problem Set 3"
author: "Ryan Shea"
date: "10/20/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(knitr)
```

I pledge my honor that I have abided by the Stevens Honor System.

\newpage

# 1: Interest Rate Derivatives

## 1.1

### (a)

```{r 1.1a}
ask <- 91 + (12.5/32)
coup_semi <- 0.02875 / 2
yield <- 0.03963
int_r_6mo <- yield / 2

previous_coup <- as.Date('2022-05-15')
today <- as.Date('2022-10-13')
next_coup <- as.Date('2022-11-15')
maturity <- as.Date('2032-05-15')

N <- round(2 * (length(integer(maturity - next_coup)) / 365))

pv_coup <- 100 * (coup_semi / int_r_6mo) * (1 - (1 / (1 + int_r_6mo)^N))

pv_ttm <- 100 / (1 + int_r_6mo)^N

ex_coup_cash <- pv_coup + pv_ttm

coup_cash <- ex_coup_cash + coup_semi


days_curr_per <- length(integer(next_coup - today))
period <- length(integer(next_coup - previous_coup))

cash_price <- coup_cash / (1 + int_r_6mo)^(days_curr_per/period)
cash_price
```

### (b)

```{r 1.1b}
cash_price_corp <- coup_cash / (1 + int_r_6mo)^((days_curr_per+1)/period)
cash_price_corp
```

\newpage

## 1.2

```{r 1.2}
px <- 4.4
basis <- 90/365


cash_px <- 100 - basis * px

(365 / 90) * log(100 / cash_px)
```


## 1.3

```{r 1.3 table, echo=FALSE}
kable(cbind(
  c("Coupon", "Maturity Date", "CUSIP", "Delivery Month 1", "Conversion Factor 1", "Delivery Month 2", "Conversion Factor 2"),
  c(4.5, "2038-05-15", "912810PX0", "December 2022", 0.8514, "March 2023", "0.8530")
))
```

```{r 1.3}
coupon <- 0.045
coup_semi <- coupon / 2
yield <- 0.06
yield_semi <- yield / 2
n <- 30

face <- (coup_semi * 100 / yield_semi) * (1 - (1 / (1 + yield_semi)^n)) + 100 / (1+yield_semi)^n

face / 100 # conversion factor

face_other <- (face + (coup_semi * 100)) / (1+yield_semi)^0.5
(face_other - (coup_semi * 100 / 2)) / 100 # Divide semi by 2 for quarterly
```

\newpage

## 1.4

```{r 1.4 table, echo=FALSE}

df <- data.frame(cbind(c(1, 2, 3, 4),
                       c(91 + (19/32), 97 + (28/32), 107 + (10/32), 116 + (18/32)),
                       c(0.8913, 0.9535, 1.0441, 1.1349))
)
colnames(df) <- c("Bond", "Quoted_Price", "Conversion_Factor")
kable(df)
```
```{r 1.4}
# df is the given table above
F_0 <- 102 + (11/32)
df$Cheapest_to_Deliver <- df$Quoted_Price - (F_0 * df$Conversion_Factor)
kable(df)
```

The cheapest to deliver is Bond 2.

\newpage

## 1.5

```{r 1.5}
today <- as.Date('2022-10-14')
delivery <- as.Date('2023-04-30')

prev_coup <- as.Date('2022-09-01')
next_coup <- as.Date('2023-03-01')

days_until_coup <- round(length(integer(next_coup - today))) # 138

days_until_delivery <- round(length(integer(delivery - today))) # 198

previous_coup_days <- round(length(integer(today - prev_coup))) # 43

days_between_coup <- round(length(integer(next_coup - prev_coup))) # 181

coup <- 0.03
coup_semi <- coup / 2
r <- 0.04

S_0 <- 85.31 + (previous_coup_days / days_between_coup) * 100 * coup_semi

S_0

# Only one coupon received during time
pv_coup <- (100 * coup_semi) * exp(-r * (days_until_coup/365))
pv_coup

F_0 <- (S_0 - pv_coup) * exp(r * (days_until_delivery / 365))
F_0

coup_future <- as.Date('2023-09-01')

sep_to_mar <- round(length(integer(coup_future - next_coup))) # 184

accr_int <- round(length(integer(delivery - next_coup))) # 60

quoted_F <- F_0 - (100 * coup_semi) * (accr_int / sep_to_mar)

# Divide by conversion factor
quoted_F / 0.6403

```

\newpage

## 1.6

### (a)

$$S_0 * e^{0.05 * 5} - S_0 * e^{(0.05 - 0.0001) * 5} = 0.1$$

$$\implies S_0 = 155.80$$


$$155.80 * e^{0.05 * 5} \approx 200.05 $$

### (b)

$$155.80 * e^{0.0499 * 5} - S_0 * e^{0.0499*10} = 0.1$$
$$\implies S_0 \approx 121.34$$

### (c)

It would most likely match **Plot 3** because the portfolio is the most optimal when as close to 5% as possible. It will begin to decrease slightly when the rates diverge away from this optimal number.

\newpage

## 1.7

### (a)

```{r 1.7a}
port_0 <- 38000000
port_duration <- 11.2 # In 4 months

F_0 <- 113
contract <- 100000
F_duration <- 17.5

N_star <- (port_0 * port_duration) / ((F_0 * contract / 100) * F_duration) 

round(N_star)
```

The optimal number of positions is to be **short 215 positions**.

### (b)

If all rates increase over the three months, but longer-term rates increase less than shorter-term rates, then there will be a gain on the short position but a loss on the actual portfolio. This is because the duration for the cheapest to deliver bonds is higher the gain on the bond portfolio.

\newpage

# 2: Swaps

## 2.1

```{r 2.1 table, echo=FALSE}
df <- data.frame(cbind(
  c(0.078, 0.082),
  c(0.084, 0.08)
))
colnames(df) <- c("US_Dollars", "AUD")
rownames(df) <- c("US_Company", "Australian_Company")

kable(df)
```
Solution on next page

![](Problem 2.1 Graph.pdf)

\newpage

## 2.2

```{r 2.2}
# pay <- 3mo LIBOR
receive <- 0.07
principal <- 50000000
# Payments every 3 months
life <- 11/12

libor_3mo <- 0.066
libor_3mo_back2mo <- 0.068
ois_r <- 0.06

val_swap <-
  principal * (receive - libor_3mo_back2mo) * 0.25 * exp(-ois_r * (1/12)) + 
  principal * (receive - libor_3mo) * 0.25 * exp(-ois_r * (4/12)) + 
  principal * (receive - libor_3mo) * 0.25 * exp(-ois_r * (7/12)) +
  principal * (receive - libor_3mo) * 0.25 * exp(-ois_r * (10/12))

val_swap
```


## 2.3

$$0 = 0.001e^{-0.03*1} + 0.001e^{-0.031 * 2} + (0.041-r)e^{-0.032*3} $$
$$\implies r \approx 0.0431$$

## 2.4

```{r 2.4}
fixed <- 0.052
principal <- 20000000
# 6 Cash flows
swap_rate <- 0.05
ois <- 0.04

val <- 0.5 * (fixed - swap_rate) * principal
total <- 0
for (i in seq(0.5, 3, 0.5)){
  total <- total + val * exp(-ois * i)
}

total

```

\newpage

## 2.5

```{r 2.5}
r_usd <- 0.051
r_gbp <- 0.048
er <- 1.12 # Exchange rate

s_usd <- 0.06 # Pays
s_gbp <- 0.06 # Receives

pr_usd <- 12000000
pr_gbp <- 10000000

V_usd <- pr_usd * s_usd * exp(-r_usd * 1) + pr_usd * (1 + s_usd) * exp(-r_usd * 2)
V_gbp <- pr_gbp * s_gbp * exp(-r_gbp * 1) + pr_gbp * (1 + s_gbp) * exp(-r_gbp * 2)

V_usd - V_gbp * 1.12
```